---
title: "Watershed 8 Test Case"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

```
Before new functionality can be added to RHESSys, new code changes must be tested for their effects on the model.    
This test case is for a small watershed (W8) located in the HJ Andrews experimental forest in Oregon.

The base simulation was run for 10 years, generating basin daily output.
Use this script to run the same simulation with your new code and compare output to this base case scenario. 

1. Edit the **rhessys_ver =** to point to your new version of RHESSys.
   (and/or edit worldfile and flowtable input files if there are changes to the structure of those files )
2. Execute the commands to run RHESSys with your new code and compare the output to this base test case output with the results in:
waterbalance_diff    
percent_diff
and various time series comparisons 
R code below guides you through this process

Download the RHESSysIOinR package at https://github.com/RHESSys/RHESSysIOinR
or directly from Rstudio with

You will also need *tidyverse* and *gpubr* that are available from R mirror sites

You don't have to run this if you have already installed RHESSYsIOinR
```{r installation}
#devtools::install_github("RHESSys/RHESSysIOinR")
```

Make sure you set your working directory to the **Testing** directory

Now insure you have needed librarier and load results from prior RHESSys simulations for comparison
```{r requiredlibraries}
#these librarier load just fine in R w/in the contianer 
library(tidyverse)
library(RHESSysIOinR)
library(ggpubr)
##Had to change this 
#load("W8TestCase.RData")
getwd() # to check where we are, now in RHESSys of docker 
# setwd("repo/Testing")
#load("W8TestCase.RData")
load("W8TestCase.RData")
out_base <- read.csv("out/base_basin_daily.csv")


```



### Input files    
Edit the **rhessys_ver** = path and file name to point to your new version of RHESSys    (in R chunk below)
Unless new development changes are being made to the worldfile or flowtable, you will use the same input files.

```{r setrunvars}

setwd("repo/Testing_rhessys_ExampleFiles")


#rhessys_ver = "/Users/naomitague/Model/RHESSys/rhessys/rhessys7.3"
#rhessys_ver = "/RHESSys/rhessys7.5" #from trials w/ Quinn
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
flowtable = "flowtables/w8TC.flow"
tecfile = "tecfiles/tec.test"

#can run 'ls()' in powershell R to see wahts in the environment 

## From 19oct trials to parse out remaking worldfiles and segementation fault issues
#rhessys_ver = "/Users/naomitague/Model/RHESSys/rhessys/rhessys7.3"
#rhessys_ver = "/RHESSys/rhessys7.5" #from trials w/ Quinn
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
flowtable = "flowtables/w8TC.flow"
tecfile = "tecfiles/tec.test" #can maybe ignore for now

#testing worldfiles I remade form their spatial data
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "w8TC_dwh4.world"
#worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
#flowtable = "flowtables/w8TC.flow"
flowtable = "w8TC_dwh4.flow"
tecfile = "tecfiles/tec.test" #can maybe ignore for now



```
#### Creates the command line
```{r commandlinegen}

cmd1 = sprintf("%s -t %s -w %s -whdr %s -r %s",
                     rhessys_ver, tecfile, worldfile, header, flowtable)
 cmd2 = sprintf("-pre out/test -s 0.355794 651.390265 -sv 0.355794 651.390265 -svalt 1.083102 1.193924, -gw 0.116316 0.916922")
#cmd2 = sprintf("-pre out/test")
cmd3 = sprintf("-st 1988 10 1 1 -ed 2000 10 1 1 -b -g")
command_line = paste(cmd1,cmd2,cmd3)
```

### Run RHESSys
```{r runrhessys}
system(command_line)
```

##Trying IOinR output filter

```{r}

setwd("repo/Testing_rhessys_ExampleFiles")


library(RHESSysIOinR)


input_rhessys = IOin_rhessys_input(
  version = "../../rhessys7.5",
  tec_file = "tecfiles/tec.test",
  world_file = "worldfiles/w8TC.world",
  world_hdr_prefix = "w8TC",
  flowtable = "flowtables/w8TC.flow",
  start = "1988 9 30 1",
  end = "2000 10 1 1",
  output_folder = "out/",
  output_prefix = "w8TC",
  commandline_options = c()
)

# Along with paths to the other basic inputs, this is the simplest way to assemble a rhessys run, though is limited in its options
run_rhessys_single(
  input_rhessys = input_rhessys,
  hdr_files = "worldfiles/w8TC.hdr",
  output_filter = "tecfiles/testing_filter.yml"
)



```




### Read in the test results and plot the water balance results

In general water balance should be close to zero - there are of course small rounding errors, typically daily water balance is less than +- 2e-06, you can see by plotting results from the water balance test for base case (included in *W8TestCase.RData*)
We will plot both so you can see if there is a water balance issue with your new code

```{r waterbalance}

getwd()
test = readin_rhessys_output("out/test")
test = readin_rhessys_output("out/3nov25_1243/test")
#DWH trials 
names(test)
test_bd <- test$bd
test_bdg <- test$bdg
# test_pdg <- test$pdg
# test_pd <- test$pd






#original code
#test$bd = watbal_basin(test$bd)

# your results are in red
#ggplot(base, aes(date, watbal))+geom_point()+geom_point(data=test$bd, aes(date, watbal), col="red", alpha=0.4)


```


#DWh plots 
basin
```{r}
test_bdg |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  select(Date, basinID, lai, surfaceN, litrc, plantc, soilc, streamflow_DON, streamflow_DOC) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  facet_wrap(~name, scales = "free_y")


test_bd |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  mutate(et = evap + trans, 
         precip_minus_et =  precip - et) |> 
  select(Date, basinID, streamflow, baseflow, lai, precip, evap, trans, et, precip_minus_et) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  #ylim(0,1)+
  facet_wrap(~name, scales = "free_y")


```

what RHESSys provides
```{r}
colnames(base)
colnames(out_base)

base |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  select(Date, basinID, streamflow) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  #ylim(0,1)+
  facet_wrap(~name, scales = "free_y")

```



patch

```{r}
unique(test_pd$patchID)
test_pd |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  select(Date, basinID, patchID, streamflow, baseflow, evap, LEcanopy)  |> 
  pivot_longer(-c(1:3)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  #ylim(0,1)+
  facet_wrap(~name, scales = "free_y")


```


### Difference between the base and the test scenarios     
The **percent_diff_mean** object shows the difference in the mean streamflow, transpiration, evaporation, psn, and plant carbon results between the base and test scenarios.     
If there are large differences - are they reasonable given the changes to functionality with your new code?

```{r percentmeandiff}
test.mean = test$bd %>% select(streamflow, psn, trans, evap, plantc) %>% summarize_if(is.numeric, mean)
base.mean =  base %>% select(streamflow, psn, trans, evap, plantc) %>% summarize_if(is.numeric, mean)

percent_diff_mean = ((test.mean-base.mean)/base.mean)*100

percent_diff_mean


base.grow.mean = base$bdg %>% select(streamflow_NO3, nitrif, denitrif, plant_resp, gpsn, soilc, litrc) %>% summarize_if(is.numeric, mean)
test.grow.mean = test$bdg %>% select(streamflow_NO3, nitrif, denitrif, plant_resp, gpsn, soilc, litrc) %>% summarize_if(is.numeric, mean)

percent_gdiff_mean = ((test.grow.mean-base.grow.mean)/base.grow.mean)*100
percent_gdiff_mean

```

## Time Series Plots

You should also look for any significant changes in key fluxes/stored
Again - if your code was designed to improve these - great - if not then make sure that changes are reasonable

First we will create some data frames 
We will then   estimate max and min differences for variables of interest
and then plot some time series

####  Time Series


```{r timeseriessummary, echo=TRUE}

# baseline 
tmp = base$bd %>% select(date, streamflow, lai, snowpack, psn, sat_def, rz_storage, trans, evap, height)
tmp2 = base$bdg %>% select(plantc, streamflow_NO3, litrc, soilc, gpsn, plant_resp, nitrif, denitrif)
tmpb = cbind.data.frame(tmp, tmp2)
tmpb$new =FALSE
# new results
tmp = test$bd %>% select(date, streamflow, lai, snowpack, psn, sat_def, rz_storage, trans, evap, height)
tmp2 = test$bdg %>% select(plantc, streamflow_NO3, litrc, soilc, gpsn, plant_resp, nitrif, denitrif)
tmpt = cbind.data.frame(tmp, tmp2)
tmpt$new =TRUE
```
Also look for big departures (needs timeseries R chunk to be run)

```{r timeseriesdep, echo=TRUE}
tmpa2 = inner_join(tmpb, tmpt, by=c("date"), suffix=c(".base",".test"))



# if any of these values are large > +- 4e-6 then you should understand the reason 
res = data.frame(matrix(ncol=0, nrow=2))

res$streamflow = c(max(tmpa2$streamflow.test-tmpa2$streamflow.base),
min(tmpa2$streamflow.test-tmpa2$streamflow.base))

res$snowpack = c(max(tmpa2$snowpack.test-tmpa2$snowpack.base),
min(tmpa2$snowpack.test-tmpa2$snowpack.base))

res$evap = c(max(tmpa2$evap.test-tmpa2$evap.base),
min(tmpa2$evap.test-tmpa2$evap.base))

res$trans = c(max(tmpa2$trans.test-tmpa2$trans.base),
min(tmpa2$trans.test-tmpa2$trans.base))

res$sat_def = c(max(tmpa2$sat_def.test-tmpa2$sat_def.base),
min(tmpa2$sat_def.test-tmpa2$sat_def.base))

res$rz_storage = c(max(tmpa2$rz_storage.test-tmpa2$rz_storage.base),
min(tmpa2$rz_storage.test-tmpa2$rz_storage.base))

res$plantc = c(max(tmpa2$plantc.test-tmpa2$plantc.base),
min(tmpa2$plantc.test-tmpa2$plantc.base))


res$soilc = c(max(tmpa2$soilc.test-tmpa2$soilc.base),
min(tmpa2$soilc.test-tmpa2$soilc.base))


res$litrc = c(max(tmpa2$litrc.test-tmpa2$litrc.base),
min(tmpa2$litrc.test-tmpa2$litrc.base))

res$psn = c(max(tmpa2$psn.test-tmpa2$psn.base),
min(tmpa2$psn.test-tmpa2$psn.base))

res$streamflow_NO3 = c(max(tmpa2$streamflow_NO3.test-tmpa2$streamflow_NO3.base),
min(tmpa2$streamflow_NO3.test-tmpa2$streamflow_NO3.base))

res$height = c(max(tmpa2$height.test-tmpa2$height.base),
min(tmpa2$height.test-tmpa2$height.base))


res=t(res)
colnames(res)=c("min","max")
res
```

If you see big departures you may want to look at plots

```{r timeseriesplots, echo=TRUE}
tmpa = rbind.data.frame(tmpb, tmpt)

a = ggplot(tmpa, aes(date, streamflow, col=new))+geom_line()+scale_y_continuous(trans="log")
b = ggplot(tmpa, aes(date, plantc, col=new))+geom_line()
c = ggplot(tmpa, aes(date, snowpack, col=new))+geom_line()
d = ggplot(tmpa, aes(date, trans, col=new))+geom_line()
e = ggplot(tmpa, aes(date, lai, col=new))+geom_line()
f = ggplot(tmpa, aes(date, evap, col=new))+geom_line()
g = ggplot(tmpa, aes(date, litrc, col=new))+geom_line()
h = ggplot(tmpa, aes(date, psn, col=new))+geom_line()
i = ggplot(tmpa, aes(date, sat_def, col=new))+geom_line()
j = ggplot(tmpa, aes(date, streamflow_NO3, col=new))+geom_line()
k = ggplot(tmpa, aes(date, soilc, col=new))+geom_line()
l = ggplot(tmpa, aes(date, height, col=new))+geom_line()
# you may have to zoom to see everything here
ggarrange(a,b,c,d,e,f,ncol=3, nrow=2)
ggarrange(g,h,i,j,k,l, ncol=3, nrow=2)
```



