---
title: "HPB trials"
output:
  html_document: default
  pdf_document: default
---

```{r installation}
#devtools::install_github("RHESSys/RHESSysIOinR")
```


```{r requiredlibraries}
#these libraries load just fine in R w/in the container 
library(tidyverse)
library(RHESSysIOinR)
library(ggpubr)

```



### Input files    
Edit the **rhessys_ver** = path and file name to point to your new version of RHESSys    (in R chunk below)
Unless new development changes are being made to the worldfile or flowtable, you will use the same input files.

Testing data
```{r}
#### TESTING trials 
setwd("repo/Testing_rhessys_ExampleFiles")

#rhessys_ver = "/Users/naomitague/Model/RHESSys/rhessys/rhessys7.3"
#rhessys_ver = "/RHESSys/rhessys7.5" #from trials w/ Quinn
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
flowtable = "flowtables/w8TC.flow"
tecfile = "tecfiles/tec.test" 

#testing worldfiles I remade form their spatial data
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "w8TC_dwh4.world"
#worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
#flowtable = "flowtables/w8TC.flow"
flowtable = "w8TC_dwh4.flow"
tecfile = "tecfiles/tec.test" 


```


HPB data
```{r setrunvars}

#can run 'ls()' in powershell R to see what is in the environment 


# ### FOR CCR
# setwd("repo")
# 
# rhessys_ver = "../rhessys7.5" #new to point backwards to model
# worldfile = "CCR_files/worldfiles/ccr.world"
# header = "CCR_files/worldfiles/ccr.hdr"
# flowtable = "CCR_files/flowtables/ccr.flow"
# tecfile = "tecfiles/tec.test" #can maybe ignore for now


### FOR HPB
#NOTE: currently using patches and worldfile based on a basin output and not patchcl, for speedier runs
setwd("repo/HPB_files_clean")

#with testing defs and veg
# rhessys_ver = "../../rhessys7.5" #new to point backwards to model
# worldfile = "worldfiles/hpb.world"
# header = "worldfiles/hpb.hdr"
# flowtable = "flowtables/hpb.flow"
# tecfile = "tecfiles/tec.test" 

setwd("repo/HPB_files_clean")

#updating veg type to decid and soil to silty loam
rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "worldfiles/hpb2.world"
header = "worldfiles/hpb.hdr"
flowtable = "flowtables/hpb2.flow"
tecfile = "tecfiles/tec.test" 




```

#### Creates the command line
Currently this is just the same from the Testing script
```{r commandlinegen}

rhessys_ver = "../../rhessys7.5" #new to point backwards to model
worldfile = "worldfiles/hpb2.world"
header = "worldfiles/hpb.hdr"
flowtable = "flowtables/hpb2.flow"
tecfile = "tecfiles/tec.test" 

cmd1 = sprintf("%s -t %s -w %s -whdr %s -r %s",
                     rhessys_ver, tecfile, worldfile, header, flowtable)
 cmd2 = sprintf("-pre out/HPBtest -s 0.355794 651.390265 -sv 0.355794 651.390265 -svalt 1.083102 1.193924, -gw 0.116316 0.916922")
 #cmd2 = sprintf("-pre out/HPBtest -s 1 2000 -sv 1 2000 -svalt 1 2000, -gw 3 3")
cmd3 = sprintf("-st 1950 10 1 1 -ed 2024 10 1 1 -b -g")
#cmd3 = sprintf("-st 1960 10 1 1 -ed 2010 10 1 1 -b -g") #DWH changed year
command_line = paste(cmd1,cmd2,cmd3)

```

### Run RHESSys
```{r runrhessys}
system(command_line)
```


### Read in the test results 


```{r}
getwd()
## inputs used for output below
#updating veg type to decid and soil to silty loam
# rhessys_ver = "../../rhessys7.5" #new to point backwards to model
# worldfile = "worldfiles/hpb2.world"
# header = "worldfiles/hpb.hdr"
# flowtable = "flowtables/hpb2.flow"
# tecfile = "tecfiles/tec.test" 

## this was originally output in repo/out, but i moved into dated folder
#hpb_test = readin_rhessys_output("../out/hpb_19oct25/test") #first run
hpb_test = readin_rhessys_output("out/3nov25_400/HPBtest")
#hpb_test = readin_rhessys_output("out/HPBtest") #most recent run that hasn't been moved to a folder

#DWH trials 
names(hpb_test)
hpb_test_bd <- hpb_test$bd
hpb_test_bdg <- hpb_test$bdg

output <- left_join(hpb_test_bd, hpb_test_bdg, by = c("day", "month", "year", "basinID", "wy", "yd", "date")) |> 
  select(day, month, year, date, yd, wy, basinID, everything())

## Figuring out what duped columns exist between outputs, was calling output 'zed' for this
# colnames(zed)
# 
# zed |> 
#   select(1:5, lai.y, lai.x, gpsn.x, gpsn.y, plantc.x, plantc.y, litrc.x, litrc.y)
#   ggplot()
#   
#   get_duplicated_stems <- function(df) {
#   # Remove .x/.y suffixes
#   base_names <- sub("\\.(x|y)$", "", colnames(df))
#   
#   # Find duplicated stems
#   dupes <- base_names[duplicated(base_names) | duplicated(base_names, fromLast = TRUE)]
#   
#   unique(dupes)
# }
# 
# # Usage
# get_duplicated_stems(zed)



```

```{r}

hpb_test_bd |> 
       mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
filter(Date > ymd("1980-01-01")) |> 
  select(Date, basinID, streamflow, baseflow, precip, recharge, unsat_stor, 
         rz_storage, sat_def, 
         litter_store, litter_capacity, 
         evap, evap_can, evap_lit, evap_soil,
         gw.Qout, gw.storage
         ) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+ 
  facet_wrap(~name, scales = "free_y")


#streamflow by year
hpb_test_bd |> 
  filter(wy > 2020) |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  select(Date, year, wy, yd, streamflow) |> 
  ggplot(aes(x = yd, y = streamflow, color = as.factor(wy) ))+
  geom_line()

  

```



Plot timeseries of outputs

```{r}

hpb_test_bdg |> 
  mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  filter(Date >= ymd("2000-10-01")) |>
  select(Date, basinID, lai, surfaceN, litrc, plantc, soilc, streamflow_DON, streamflow_DOC) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  facet_wrap(~name, scales = "free_y")

colnames(hpb_test_bd)
colnames(hpb_test_bdg)

hpb_test_bd |> 
   mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
  filter(Date >= ymd("2000-10-01")) |>
  mutate(et = evap + trans, 
         precip_minus_et =  precip - et) |> 
  select(Date, basinID, streamflow, baseflow, precip, evap, lai, trans, et, precip_minus_et) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  #ylim(0,1)+
  facet_wrap(~name, scales = "free_y")

hpb_test_bd |> 
   mutate(Date = ymd(paste(year, month, day, sep = "-"))) |> 
    # filter(Date >= ymd("1990-10-01"),
    #      Date <= ymd("2010-10-01")) |>
  select(Date, basinID, streamflow, baseflow, precip, recharge, unsat_stor, 
         rz_storage, sat_def, 
         litter_store, litter_capacity, 
         evap, evap_can, evap_lit, evap_soil,
         gw.Qout, gw.storage
         ) |> 
  pivot_longer(-c(1:2)) |> 
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  #ylim(0,1)+
  facet_wrap(~name, scales = "free_y")

```



```{r}

# #read in original climate files as a check
# getwd()
# 
# ccrclim <- read_clim("clim/ccr_daily")
# 
# ccrclim |> 
#   mutate(Date = ymd(date)) |> 
#   filter(Date >= ymd("1989-10-01"),
#          Date <= ymd("2000-09-30")) |> 
#   select(Date, rain, tmin, tmax) |> 
#   pivot_longer(-1)  |> 
#   ggplot(aes(x = Date, y = value))+
#   geom_point()+
#   facet_wrap(~name, scales = "free_y")
```

